---
title: "Query Histocial Records from Columbia University History Lab - Declassification Engine API"
author: "Jianghanhan Li"
date: "2017-12-15"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# R Package for History Lab Declassification Engine API 

Declassification Engine API is provided by [Columbia University History Lab](http://www.history-lab.org/). 

#### Setup

```{r}
devtools::install_github("hadley/devtools")
library(devtools)
library(roxygen2)
library(testthat)
```

```{r}
library(httr)
library(jsonlite)
library(rvest)
```

#### Dispaly Welcome Page 

Use `declass_welcome` function to see the welcome page of Declassification Engine API. Currently it just shows "Welcome to the Declassification Engine REST API". The History Lab might add more content in the future. 

```{r}
declass_welcome <- function(return_type = "string"){
  url <- "http://api.declassification-engine.org/declass"
  message <- httr::content(httr::GET(url))
  if (return_type == "html") {
    return(message)
  } else if (return_type == "string") {
    text <- rvest::html_text(message)
    return(text)
  } else {
    notice <- "Return_type not supported"
    return(notice)
  }
}
```

You can see the welcome page displayed in different data types. 

```{r}
test_declass_welcome <- function() {
  print(declass_welcome("html"))  
  print(declass_welcome("string"))
  print(declass_welcome("xml"))
  print(declass_welcome())
}
test_declass_welcome()
```


#### Display all available document metadata fields 

Use `declass_fields` function to see all the metadata fields. Towards the end, there're many repetitive fields. 

```{r}
declass_fields <- function(return_type = "list") {
  url <- "http://api.declassification-engine.org/declass/v0.4/fields"
  if(return_type == "list") {
    parsed_json <- jsonlite::fromJSON(url)
    fields <- parsed_json$results$fields
    return(fields)
  } else if(return_type == "raw") {
    raw <- content(GET(url))
    return(raw)
  } else {
    notice <- "Return_type not supported"
    return(notice)
  }
}
```

Similarly, see the metadata fields in different data formats, depending on your need. 

```{r}
test_declass_fields <- function() {
  print(declass_fields("list"))
  print(declass_fields("raw"))
  print(declass_fields())
}
test_declass_fields()
```


#### List all available collections 

Use `declass_collection` to see the name of all collections. 

```{r}
declass_collection <- function(return_type = "list"){
  url <- "http://api.declassification-engine.org/declass/v0.4/collections"
  parsed_json <- jsonlite::fromJSON(url)
  collections <- parsed_json$results$collections
  if (return_type == "list") {
    return(collections)
  } else if (return_type == "raw") {
    raw <- content(GET(url))
    return(raw)
  } else {
    notice <- "Return_type not supported"
    return(notice)
  }
}
```

Check it out, as the content displayed in different formats. If you didn't specify return type, a message will appear.

```{r}
test_declass_collection <- function() {
  print(declass_collection("list"))  
  print(declass_collection("raw"))
  print(declass_collection())
}

test_declass_collection()
```


#### Display all the entities for a particular collection 

Use `declass_collection_entity` to see specific entites in a specific collection. 

```{r}
declass_collection_entity <- function(collection) {
  if (collection %in% c("cpdoc", "clinton", "kissinger", "statedeptcables", "frus", 
                        "ddrs", "cabinet", "pdb")) {
    url <- paste("http://api.declassification-engine.org/declass/v0.4/entity_info/?collection=", 
                 collection, sep = "")
    parsed_json <- jsonlite::fromJSON(url)
    collection <- parsed_json$results$entity
    return(collection)
  } else {
    notice <- "Collection name and entities not found"
    return(notice)
  }
}

```

```{r}
test_declass_collection_entity <- function() {
  print(declass_collection_entity("cpdoc"))  
  print(declass_collection_entity("clinton"))
  print(declass_collection_entity("kissinger"))
  print(declass_collection_entity("statedeptcables"))
  print(declass_collection_entity("frus"))
  print(declass_collection_entity("ddrs"))
  print(declass_collection_entity("cabinet"))
  print(declass_collection_entity("pdb"))
  print(declass_collection_entity(""))
}

test_declass_collection_entity()
```


#### Query data for a specified entity in a collection 

Use `declass_entity_data` to query entity data for a specific collection. You can customize the `page_size` which shows how many records to be displayed per page, `page` which specifies the starting page, `num_pages` which specifies the number of pages to be queried. 

```{r}
declass_entity_data <- function(collection, entity, page_size = 25, page = 1, num_pages = 10) {
  url <- paste("http://api.declassification-engine.org/declass/v0.4/entity_info/?collection=",
               collection, "&entity=", entity, "&page_size=", page_size, "&page=", page, sep = "")
  
   if (collection %in% c("cpdoc",  "kissinger", "cabinet") & 
       entity %in% c("countries", "persons", "topics")) {
    }
   else if (collection %in% c("statedeptcables", "frus", "ddrs", "pdb") & 
            entity %in% c("countries", "persons", "topics", "classifications")) {
    }
   else if (collection == "clinton" & entity %in% c("countries", "persons", "classifications")) {
      } 
  else {
    notice <- "Collection or entity name not found"
    return(notice)
    
  }
  print(url)
  parsed_json <- jsonlite::fromJSON(url)
  entity_data <- parsed_json$results
  num_pages <- num_pages - 1

  while (is.null(parsed_json$next_page) == FALSE & num_pages > 0) {
    print(parsed_json$next_page)
    parsed_json <- jsonlite::fromJSON(parsed_json$next_page)
    entity_data <- rbind(entity_data, parsed_json$results)
    num_pages <- num_pages - 1
  }
  return(entity_data)
}

```

```{r}
test_declass_entity_data <- function() {
print(declass_entity_data("kissinger", "persons", page_size = 1, num_pages = 5))
print(declass_entity_data("frus", "topics"), page_size = 20, num_pages = 3)
print(declass_entity_data("frus", "countries", page_size = 50, num_pages = 1))
print(declass_entity_data("clinton", "topics"))
}

test_declass_entity_data()
```

Here the topics in these collections were re-engineered using Topic Modeling, as exactly the same as available in the Declassification Engine API. 

#### Return random document ids with a limit 

Use `declass_query_random_ids` to obtain randomly sampled IDs to play around and view documents for lay users. 

```{r}
declass_query_random_ids <- function(limit = 10) {
  base_url <- "http://api.declassification-engine.org/declass/v0.4/random/?limit="
  if (limit <= 1024) {
    url <- paste(base_url, limit, sep = "")
    ids <- jsonlite::fromJSON(url)$results
    return(ids)
  } else {
    notice <- "Limit exceeds the maximum 1024"
    return(notice)
  }
}
```

```{r}
test_declass_query_random_ids <- function() {
print(declass_query_random_ids(32))
print(declass_query_random_ids(1100))
}
test_declass_query_random_ids()
```


#### Query document by ID 

Use `declass_query_id_data` to manually input ID or list of IDs to view documents.

```{r}
declass_query_id_data <- function(id_list) {
  base_url <- "http://api.declassification-engine.org/declass/v0.4/?ids="
  url <- paste(base_url,id_list, sep = "")
  doc <- jsonlite::fromJSON(url)
  return(doc)
}

```

```{r}
id <- c("Clinton-74931")
declass_query_id_data(id)
```


#### Query documents based on randomly sampled ids 

Use `declass_query_random_ids_data` to view documents using randomly sampled IDs, a function that's built on topic of the previous two functions. 

```{r}
declass_query_random_ids_data <- function(limit) {
  ids <- declass_query_random_ids(limit)
  id_list <- paste(ids$id, collapse = ",")
  id_list <- URLencode(id_list)
  return (declass_query_id_data(id_list))
}
```

```{r}
declass_query_random_ids_data(5)
```

